---
title: "Project 1"
author: Muhammad Ahsan
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---


## Cardio Fitness Project

### 1. Project Objective
The  objective  of  the  report  is  to  explore  the  cardio  data  set  (“CardioGoodFitness”)  in  R  and generate insights about the data set. 
This exploration report will consists of the following: 
 
  1. Importing the dataset in R      
  2. Understanding the structure of dataset      
  3. Graphical exploration      
  4. Descriptive statistics      
  5. Insights from the dataset 

### 2. Assumptions






### 3.  Exploratory Data Analysis – Step by step approach
A Typical Data exploration activity consists of the following steps: 
 
#### 3.1 Environment Set up and Data Import 

###### Required Libraries 

```{r results='hide', message=FALSE, warning=FALSE}
##library(readr)
library(ggplot2)
library(gridExtra)
library(corrplot)
```


##### Setting the working Directory and importing the Dataset
In this chunk we set the working directory and importing the dataset for analysis and we convert the dataset in Data Frames

```{r message = FALSE}
setwd("F:/DSBA/Project 1")
cardio_data <- read.csv("CardioGoodFitness.csv")
cardio_data <- data.frame(cardio_data)
```


#### 3.2 Variable Identification
we are using "dim()" function to get the dimention or shape of the dataset.
"names()" this function we going to use is getting the names or columns of the dataset.
"head()" this function is use for presenting top some rows or observations of the dataset bydefault we ill get top 5 rows but we can add specific also."
"tail()" retun the last 5 rows by default

dimention of the dataset
```{r}
dim(cardio_data)
```

Columns of the dataset
```{r}
names(cardio_data)
```
Top 5 rows of the dataset
```{r}
head(cardio_data)
```
Last 5 rows of the dataset
```{r}
tail(cardio_data)
```
Convertion of some columns datatype cracter to factor
```{r}
cardio_data$Product <- as.factor(cardio_data$Product)
cardio_data$Gender <- as.factor(cardio_data$Gender)
cardio_data$MaritalStatus <- as.factor(cardio_data$MaritalStatus)
cardio_data$Education <- as.factor(cardio_data$Education)
```

Summary of the data 
```{r}
#Statisticl Summary of Dataset
summary(cardio_data)
```
### 3.3  Univariate Analysis

You can also embed plots, for example:
```{r}
#Statisticl Summary of Product 
summary(cardio_data$Product)
#product plot 
p<- ggplot(data = cardio_data, aes(x = Product)) + 
  geom_bar(color="black" ,fill="darkcyan") +
  theme_grey() + ggtitle("Product Plot")+xlab("Product")+
  geom_text(stat = 'count',aes(label =..count.., vjust = -0.2))
p
#p+geom_text()
    
```

```{r}
#Statisticl Summary of Age
summary(cardio_data$Age)
#ge plot 
p1<- ggplot(data = cardio_data, aes(x = Age)) + 
  geom_histogram(bins = 30, color="black" ,fill="darkcyan") +
  theme_grey() + ggtitle("Histogram of Age")+xlab("Age (Years)")
#Statisticl Summary of education
summary(cardio_data$Education)
#education plot 
p2<-ggplot(data = cardio_data, aes(x = Education)) + 
  geom_bar(color="black" ,fill="darkcyan") +
  theme_grey() + ggtitle(" Education Bar plot ")+
  geom_text(stat = 'count',aes(label =..count.., vjust = -0.2))
#Statisticl Summary of usage
summary(cardio_data$Usage)
#usage plot 
p3<-ggplot(data = cardio_data, aes(x = Usage)) + 
  geom_bar(color="black" ,fill="darkcyan") +
  theme_grey() + ggtitle(" Usage of Machines ")+
  geom_text(stat = 'count',aes(label =..count.., vjust = -0.2))
#combining the plots
grid.arrange(p1,p2,p3,ncol=3)
```

```{r}
#Statisticl Summary of Gender
summary(cardio_data$Gender)
ggplot(data = cardio_data, aes(x = Gender)) + 
  geom_bar(color="black" ,fill="darkcyan") +
  theme_grey() + ggtitle("Gender Ploting")+
  geom_text(stat = 'count',aes(label =..count.., vjust = -0.2))
```

```{r}
```

```{r}
#Statisticl Summary of Marital Status
summary(cardio_data$MaritalStatus)
#Maritel status plot 
p1<- ggplot(data = cardio_data, aes(x = MaritalStatus)) + 
  geom_bar( color="black" ,fill="darkcyan") +
  theme_grey() + ggtitle("User Marital Status")+
  geom_text(stat = 'count',aes(label =..count.., vjust = -0.2))
#Statisticl Summary of income
summary(cardio_data$Income)
#Income plot
p2<- ggplot(data = cardio_data, aes(x = Income)) + 
  geom_histogram(bins = 50, color="black" ,fill="darkcyan") +
  theme_grey() + scale_x_continuous(labels = scales::comma) + ggtitle("Histogram Of User Income")


p3<- ggplot(data = cardio_data, aes(x = "",y=Income)) + 
  geom_boxplot(color="black" ,fill="darkcyan") +
  theme_grey() +  ggtitle("User Income")+ scale_y_continuous(labels = scales::comma) 
#combining the plots
grid.arrange(p1,p2,p3,ncol=3)
```

### 3.4  bivariate Analysis

```{r}
#statistical summary of Age and Gender
tapply(cardio_data$Age,cardio_data$Gender,summary)
ggplot(cardio_data,aes(x= Gender,y=Age)) +geom_boxplot(aes(colour = Gender), outlier.colour = "red",notch=TRUE) 
```

```{r}
#statistical summary of Age and Gender
tapply(cardio_data$Age,cardio_data$Gender,summary)
ggplot(cardio_data,aes(x= Gender,y=Age)) +geom_boxplot(aes(colour = Education), outlier.colour = "red",notch=FALSE) 

```

```{r}
#statistical summary of Age and Gender
tapply(cardio_data$Age,cardio_data$Product,summary)
ggplot(cardio_data,aes(x= Product,y=Age)) +geom_boxplot(aes(colour = Product), outlier.colour = "red",notch=FALSE) 

```

```{r}
#statistical summary of Age and Gender
tapply(cardio_data$Fitness,cardio_data$Product,summary)
ggplot(cardio_data,aes(x= Product,y=Fitness)) +geom_point(aes(colour = Product), outlier.colour = "red",notch=FALSE) 

```

```{r}
#statistical summary of Age and Gender
tapply(cardio_data$Income,cardio_data$Product,summary)
ggplot(cardio_data,aes(x= Product,y=Income)) +geom_boxplot(aes(colour = Product), outlier.colour = "red",notch=FALSE) + scale_y_continuous(labels = scales::comma)

```

```{r}
#statistical summary of Age and Gender
tapply(cardio_data$Miles,cardio_data$Gender,summary)
ggplot(cardio_data,aes(x=Gender ,y=Miles)) +geom_boxplot(aes(fill=Gender), outlier.colour = "red",notch=FALSE) + scale_y_continuous(labels = scales::comma)
```

```{r}
#statistical summary of Age and Gender
ggplot(cardio_data,aes(x=factor(Usage) ,y=Income)) +geom_boxplot(aes(fill=Usage), outlier.colour = "red",notch=FALSE) + scale_y_continuous(labels = scales::comma)
```

```{r}
#statistical summary of Age and Gender
  tapply(cardio_data$Fitness,cardio_data$Gender,summary)
ggplot(cardio_data,aes(x= Gender,y=Fitness)) +geom_boxplot(aes(fill=Gender), outlier.colour = "red",notch=FALSE) 
```

```{r}
corrplot(cor(cardio_data[,6:9]))
```

```{r}
#statistical summary of Age and Gender
tapply(cardio_data$Miles,cardio_data$Fitness,summary)
ggplot(cardio_data,aes(x=factor(Fitness),y=Miles)) +
  geom_boxplot(aes(fill=MaritalStatus), outlier.colour = "red",notch=FALSE) 
```

```{r}

```